Quantitative Auditing of Fairness Measures with Differentially Private Synthetic Data

### 2024-12-01 paper/paper.tex
alot of progress

\clearpage

@@

@@

Let $C \subseteq \mathcal{A}$ be a subset of attributes. Let $\Omega_C = \prod_{i \in C} \Omega_i$. Let $x$ be a row and $x_C$ denote the restriction of $x$ to $C$. A \emph{marginal}\cite{barak2007privacy,mckenna2021winning} of $C$ on database $\db$ is a lookup table $\mu_{\db} : \Omega_C \implies \mathbb{N}_0$ such that each entry is a count $\mu_{\db}(t) = \Sigma_{x \in \db} \delta_{t,{x_C}}$ where $\delta$ is the Kronecker function; that is, it is the vector of the counts of each possible domain element. We call marginals of $|C| = n$ attributes $n$-way marginals.

The task of differentially private synthetic data\cite{Ullman2022,McKenna2022,NIST_Differentially_Private_Synthetic_Data,YouTube_Differentially_Private_Synthetic_Data} is, given a database $\db$, adding some noise to marginals of $\db$ such that it satisfies some differential privacy guarantees and outputting another database $\db'$, such that the $L^1$-norm errors between marginals of $\db$ and $\db'$ is small; that is, their marginals $\mu_{\db},\mu_{\db'}$ are similar.

@@

The synthesis framework is three-fold, namely, select-measure-generate\cite{McKenna2022}. We first select the important marginals to preserve, measure them by adding differentially private noise, and then generate synthetic data.

@@

The fairness checker is an open-sourced public domain Python package that computes various fairness measures, such as those mentioned in Table~\ref{measures}.

>>>>

Fairness in machine learning has become a crucial area of research\cite{barocas2023fairness}. Various fairness measures have been proposed to quantify the fairness of AI models\cite{pessach2022review,corbett2017algorithmic,vzliobaite2017measuring,hardt2016equality,corbett2017algorithmic,berk2021fairness,chouldechova2017fair,kleinberg2016inherent}.

The auditing of AI models is an important step in ensuring that these models are fair\cite{ferrara2023fairness}. Different tools have been developed to check the fairness of AI models\cite{yuan2024ensuring,saleiro2018aequitas,bellamy2019ai,bird2020fairlearn}.

% differential privacy

Differential Privacy has risen to become the standard of privacy protection in many data analyses\cite{jiang2021applications}. It offers a formal framework for quantifying privacy guarantees when releasing information derived from sensitive data\cite{dwork2006calibrating,dwork2014algorithmic}.

Different flavors of Differential Privacy have been proposed over the years\cite{desfontaines2019sok}, such as Gaussian Differential Privacy\cite{dong2022gaussian}, Pufferfish Differential Privacy\cite{kifer2012rigorous}, Bayesian Differential Privacy\cite{triastcyn2020bayesian}, and R\'enyi Differential Privacy\cite{mironov2017renyi}.

% synthetic data

The generation of synthetic data under differential privacy constraints allows for data sharing without compromising privacy\cite{tao2021benchmarking}. There have been several techniques developed for Differentially Private synthetic data generation\cite{rosenblatt2020differentially,fan2020relational,bowen2019comparative,bowen2020comparative,arnold2020really,xu2019modeling} in recent years.

While useful, there are some ethical concerns with the generation of synthetic data ...

% TODO: cite

@@

% TODO: should we exclude nonbinary prediction?

@@

Let $C \subseteq \mathcal{A}$ be a subset of attributes. Let $\Omega_C = \prod_{i \in C} \Omega_i$. Let $x$ be a row and $x_C$ denote the restriction of $x$ to $C$. A \emph{marginal}\cite{barak2007privacy,mckenna2021winning} of $C$ on database $\db$ is a lookup table $\mu_{\db} : \Omega_C \implies \mathbb{N}_0$ such that each entry is a count $\mu_{\db}(t) = \Sigma_{x \in \db} \delta_{t,{x_C}}$ where $\delta$ is the Kronecker function; that is, it is a table of the counts of each possible combination of domain elements. We call marginals of $|C| = n$ attributes $n$-way marginals.

The task of differentially private synthetic data\cite{Ullman2022,McKenna2022,NIST_Differentially_Private_Synthetic_Data,YouTube_Differentially_Private_Synthetic_Data} is, given a database $\db$, adding some noise to marginals of $\db$ such that it satisfies some Differential Privacy guarantees and outputting another database $\db'$, such that the $L^1$-norm errors between marginals of $\db$ and $\db'$ is small; that is, their marginals $\mu_{\db},\mu_{\db'}$ are similar.

@@

The synthesis framework is three-fold, namely, select-measure-generate\cite{McKenna2022}. We first select the important marginals to preserve, measure them by adding Differential Privacy noise, and then generate synthetic data.

@@

The fairness checker is an open-sourced public domain Python package that computes various fairness measures, such as those mentioned in Table~\ref{tab:measures}.
